Malacology: A Programmable Storage System

### 2016-10-18 paper/paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

\paragraph*{Malacology:} interfaces are added in strategic spots in the
metadata service for guarding shared resources and load balancing.

@@

{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
metadata load balancer, has been implemented on Malacology so that its
infrastructure can enjoy all the properties of existing internal abstractions.
The load balancing logic is intact but the infrastructure has been improved to
safely load, manage, and persist its load balancing policies.  Composing Mantle
using the service metadata and durability interfaces, Mantle can safely
version balancer policies, save balancer policies in the backend object store
and centralize warnings/errors.

@@

The second service implemented on Malacology is ZLog, a high-performance
distributed shared-log that is based on the CORFU
protocol~\cite{balakrishnan_corfu_2012}. A shared-log is a general, yet
powerful abstraction used to build many distributed systems, such as
cloud-based metadata management~\cite{balakrishnan:sosp13} and elastic database
storage engines~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
However, existing implementations that rely on consensus algorithms such as
Paxos funnel I/O through a single point introducing a bottleneck that restricts
throughput.  In contrast, the CORFU protocol is able to achieve high throughput
using a network counter service, called a \emph{sequencer}, for obtaining the
entry at the tail of log. This metadata operation that avoids all persistent
I/O in the common case.

@@

\subsubsection{Sequencer}

@@

We implement the sequencer service in Malacology as an application-specific
file type that associates a small amount of metadata, including the 64-bit log
tail position, to the inode state managed by the Ceph metadata service. Since
a sequencer is associated with a unique log, this approach has the added
benefit of allowing the metadata service to also handle naming, by
representing each log service instance within the standard POSIX hierarchical
namespace.

{\bf Sequencer interface.} The sequencer resource supports the ability to
\emph{read()} the current tail value as well get the \emph{next()} position in
the log which also atomically increments the tail position. We implement this
functionality as methods on ZLog inode resources. The primary challenge in
mapping the sequencer resource to the metadata service is handling
serialization correctly.

Initially we sought to directly model the sequencer service in Ceph as a
non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
access to the resource at the authoritative metadata server for the sequencer
inode.  Interestingly, we found that the capability system in Ceph uses a
strategy to reduce metadata service load by allowing multiple clients that
open a shared file to temporarily cache the inode resource, resulting in a
round-robin, best-effort batching behavior. When a single client is accessing
the sequencer resource it is able to cache state locally without interruption.

@@

The storage interface is a critical component in the CORFU design. Clients
independently map log positions that they have obtained from a sequencer
service (described in detail in the next section) onto storage devices, while
storage devices provide an intelligent \emph{write}-once, random \emph{read}
interface for accessing log entries. The key to correctness in CORFU lies with
the enforcement of up-to-date epoch tags on client requests; requests tagged
with out-of-date epoch values are rejected, and clients are expected to
request a new tail from the sequencer after refreshing state from an auxiliary
service.  This mechanism forms the basis for sequencer recovery.

In order to repopulate the cached tail value during recovery of a sequencer,
the maximum position in the log must be obtained. To do this, the storage
interface exposes an additional \emph{seal} method that atomically installs a
new epoch value and returns the maximum log position that has been written.
Since the recovery protocol does not inform clients of a new epoch value until
the end of the process, the validity of the tail can be guaranteed despite the
iterative process of contacting storage devices.  Recovery of the sequencer
process itself may be handled in many ways, such as leader election using an
auxiliary service such as Paxos. In our implementation, the recovery is the
same (and is inherited from) the Ceph metadata service.

>>>>

\paragraph*{Malacology:} interfaces are added in strategic spots in the
metadata service for guarding shared resources, defining new file types,
and load balancing.

@@

{\bf File types.} The metadata service manages a POSIX file system hierarchy
where files and directories are represented as inode data structures and
expose a POSIX file interface. Application that manage large amounts of
metadata (e.g. users or database snapshots) often require a naming service. We
allow new inode types to be defined such that applications can create
domain-specific interfaces to inodes that may modify locking and capability
policies. We will show this is used in Section~\ref{sec:seq} when we discuss a
distributed shared-log built on Malacology.

{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
metadata load balancer, has been implemented on Malacology so that its
infrastructure can enjoy all the properties of existing internal abstractions.
The load balancing logic is intact but the infrastructure has been improved to
safely load, manage, and persist its load balancing policies.  Composing Mantle
using the service metadata and durability interfaces, Mantle can safely
version balancer policies, save balancer policies in the backend object store
and centralize warnings/errors. When combined with file types, load balancing
policies can express a variety of rules such as multi-tentant and
multi-application balancing.

@@

The second service implemented on Malacology is ZLog, a high-performance
distributed shared-log that is based on the CORFU
protocol~\cite{balakrishnan_corfu_2012}. A shared-log is a general, yet
powerful abstraction used to construct distributed systems, such as
cloud-based metadata management~\cite{balakrishnan:sosp13} and elastic database
storage engines~\cite{bernstein:cidr11,bernstein:vldb11,bernstein:sigmod15}.
However, existing implementations that rely on consensus algorithms such as
Paxos funnel I/O through a single point introducing a bottleneck that restricts
throughput.  In contrast, the CORFU protocol is able to achieve high throughput
using a network counter service, called a \emph{sequencer}, that is used to
decouple log position assignment from log I/O.

@@

\subsubsection{Sequencer}
\label{sec:seq}

@@

{\bf Sequencer interface.} The sequencer resource supports the ability to
\emph{read()} the current tail value as well as get the \emph{next()} position in
the log which also atomically increments the tail position.
We implement the sequencer service in Malacology as an application-specific
file type that associates a small amount of metadata, including the 64-bit log
tail position, to the inode state managed by the Ceph metadata service. Since
a sequencer is associated with a unique log, this approach has the added
benefit of allowing the metadata service to also handle naming, by
representing each log service instance within the standard POSIX hierarchical
namespace. The primary challenge in mapping the sequencer resource to the
metadata service is handling serialization correctly to maintain the global
ordering provided by the CORFU protocol.

Initially we sought to directly model the sequencer service in Ceph as a
non-exclusive, non-cacheable resource, forcing clients to perform a round-trip
access to the resource at the authoritative metadata server for the sequencer
inode.  Interestingly, we found that the capability system in Ceph uses a
strategy to reduce metadata service load by allowing clients that open a
shared file to temporarily obtain an exclusive cached copy of the resource,
resulting in a round-robin, best-effort batching behavior. When a single
client is accessing the sequencer resource it is able to increment the
sequencer locally. Any competing client cannot query the sequencer until the
metadata service has granted it exclusive access.

@@

The storage interface is a critical component in the CORFU protocol. Clients
independently map log positions that they have obtained from the sequencer
service (described in detail in the next section) onto storage devices, while
storage devices provide an intelligent \emph{write}-once, random \emph{read}
interface for accessing log entries. The key to correctness in CORFU lies with
the enforcement of up-to-date epoch tags on client requests; requests tagged
with out-of-date epoch values are rejected, and clients are expected to
request a new tail from the sequencer after refreshing state from an auxiliary
service.  This mechanism forms the basis for sequencer recovery.

In order to repopulate the sequencer state (i.e. the cached, current tail of
the log) during recovery of a sequencer,
the maximum position in the log must be obtained. To do this, the storage
interface exposes an additional \emph{seal} method that atomically installs a
new epoch value and returns the maximum log position that has been written.

Since the sequencer servivce does not resume until the recovery process has
completed, there cannot be a race with clients appending to the log, and the
immutability of the log allow reads to never block during a sequencer failure.
Recovery of the sequencer process itself may be handled in many ways, such as
leader election using an auxiliary service such as Paxos. In our
implementation, the recovery is the same (and is inherited from) the Ceph
metadata service. Handling the failure of a client that holds the sequencer
state is similar, although a timeout is used to determine when a client should
be considered unavailable.
